# Analysis

## Descriptive statistics

Simple summary statistics can sometimes be obtained with `summary()` -- this will give you min, max, and mean for your dependent variables, which might be enough for your purposes.

* `summary()`

## Aggregation

Once you have a dataset containing results from all participants, you might want to aggregate over trials, participants and conditions before you proceed to analysis.

Aggregation could be a step that you perform in its own right, to obtain descriptive statistics at the condition level. In this case, `summary()` won't meet your needs because you need to aggregate over trials in each condition, per participant, to obtain "participant means", and then go on to aggregate over participant means to the condition level to get "condition means".

Aggregation could be a step that you need to perform before doing inferential statistics. In this case it's likely that you want to aggregate over trials in each condition, per participant, to obtain "participant means", and submit the participant means to analysis.

In either case, aggregation is done by:

* `group_by()`
* `summarise()`

## Example with Code

* **raw data**

```{r, echo=F}
set.seed(123)
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(kableExtra))
suppressPackageStartupMessages(library(Hmisc))
```

```{r, echo=T, message=F}
set.seed(2)
raw_data <-
  tibble(
    participant = as_factor(rep(c("Adam", "Bob", "Charlie"), each=4)), 
    condition = as_factor(rep(c("drug", "placebo"), each=2, times=3)),
    trial = rep(c(1,2), times=6),
    score = sample(25:50, 12, replace=F)
    ) %>% 
  mutate(score=ifelse(condition=="drug", score+20, score))
```

```{r, echo=F}
raw_data 
```

* **participant means**: a single value per-condition for each participant, collapsing over trials

```{r, message=F}
participant_means = raw_data %>% 
  group_by(participant, condition) %>% 
  summarise(mean_score=mean(score))
```

```{r, echo=F}
participant_means
```

* **condition means**: a single value for each condition, collapsing over participants

```{r, message=F}
condition_means = participant_means %>% 
  group_by(condition) %>% 
  summarise(grand_mean_score=mean(mean_score))
```

```{r, echo=F}
condition_means
```

## Inferential statistics

```{r}
library(ez)
```

```{r}
ANOVA_OUTPUT = 
  ezANOVA(
    data = participant_means,
    dv = mean_score,
    wid = participant,
    within = condition,
    return_aov = TRUE
  )
```
```{r}
ANOVA_OUTPUT$ANOVA %>% select(-ges) %>% kable(row.names = F, digits=4)
```

ygrvgr iursvgiuewqviurewuvi 


```{r}
ggplot(data=participant_means, aes(y=mean_score, x=condition, color=condition))+
  #geom_violin()+
  stat_summary(fun.data=mean_cl_normal)+
  theme_bw(base_size=14)+
  theme(panel.grid=element_blank())
```